Toward Adaptive GPU Resource Management for Embedded Real-Time Systems

### 2012-03-09 draft/draft.tex
Abstract is added, and the introduction is modified accordingly.

\begin{abstract}
 In this paper, we present conceptual frameworks for GPU applications to
 adjust their task execution times upon given workloads.
\end{abstract}

@@

In this paper, we propose conceptual frameworks for GPU applications to adjust their task execution times depending on their given workloads.

@@

Our second approach to adaptive resource management is an implicit
adjustment framework.
In this framework, the number of cores to be allocated for the program
is set by the runtime system.
Hence, adaptivity of computing does not really depend on programmers,
but if the program is not aware of this framework, it may fail to run,
since the number of core allocated for the program may be different from
what the program assumes.

@@

>>>>

\begin{abstract}
 In this paper, we present conceptual frameworks for GPU applications to
 adjust their task execution times upon given workloads. We propose two 
 frameworks enabling smart GPU resource management when many applications 
 share GPU resources while the workloads of those applications vary. 
 For {\it explicit adjustment}, we provide API functions so that application 
 developers can adjust the number of GPU cores depending on their needs. 
 An {\it implicit adjustment} will provide a run-time framwork which 
 dynamically allocates the number of cores to tasks based on the given 
 workloads. The runtime support of the proposed system can be realized using 
 functions which measure the execution times of the tasks on GPU and change 
 the number of GPU cores. We motivate the necessity of this framework 
 in the context of self-driving technologies, and we believe that our 
 frameworks for GPU programming are useful contributions given that 
 trends are apparently shifting to highly parallel heterogeneous computing.
\end{abstract}

@@

In this paper, we present conceptual frameworks for
GPU applications to adjust their task execution times upon
given workloads. We propose two frameworks enabling
smart GPU resource management when many applications
share GPU resources while the workloads of those appli-
cations vary. For explicit adjustment, we provide API func-
tions so that application developers can adjust the number
of GPU cores depending on their needs. An implicit adjust-
ment will provide a run-time framwork which dynamically
allocates the number of cores to tasks based on the given
workloads. The runtime support of the proposed system
can be realized using functions which measure the execu-
tion times of the tasks on GPU and change the number of
cores.

@@

Our second approach to adaptive resource management is an implicit
adjustment framework.
In this framework, the number of cores to be allocated for the program
is set by the runtime system.
Hence, adaptivity of computing does not really depend on programmers.
If the program is not aware of this framework, however, it may fail to
run, since the number of core allocated for the program may be different
from what the program assumes.

@@

\subsection{Runtime System Support}
\label{sec:runtime}

The runtime system provides the API for programmers.
In order to support adaptive GPU resource management, we must provide
some additional API functions.

\begin{itemize}
 \item Our adaptive GPU resource management frameworks require a
       function to measure the execution time of each job running on the
       GPU.
       This function is easy-to-implement.
       Since we assume that job execution on the GPU is non-preemptive,
       the amount of time in run-to-completion can be accounted as a job
       execution time.
       This accounting method is known to work in previous
       work~\cite{Kato_ATC11, Rossbach_SOSP11}, and hence we can adopt
       it.
       For the explicit adjustment framework, this function must be
       exposed to the programmer, while it can also be used internally
       by the implicit adjustment framework.
 \item We also need several functions to change the number of cores to
       be allocated for the program.
       Some existing programming languages for GPGPU, \textit{e.g.},
       CUDA~\cite{CUDA}, provide the API to allow the programmer to
       specify the shape of grid structure and the number of threads
       mapped to compute cores.
       We can use this API as it is, or provide a corresponding API if
       the underlying programming language does not support it.
\end{itemize}

In addition to these API functions, the runtime system must be able to
detect when the program is offloaded onto the GPU and when it is
completed on the GPU.
Since the programmer calls a specific API function to launch the GPU
program in most GPU programming models, it is very easy to record the
start time of GPU execution.
Detection of the completion time of GPU execution, on the other hand, is
not straightforward.
We would need to use an interrupt to notify the runtime system of
completion of GPU execution.
Polling on a particular register is an alternative approach, but it
would not be suitable for latency-sensitive real-time systems, as
previous work demonstrated~\cite{Kato_ATC11}.

Finally, runtime system support must be integrated with the API so that
the programmer can make use of our frameworks under a single unified
programming model.
We plan to extend our CUDA runtime library developed in previous
work~\cite{Kato_OSPERT11} to support our adaptive frameworks.
This is our prototype implementation, and our frameworks can also be
integrated with other programming models beyond CUDA.
